Reproducible development environments with rix

### 2025-02-18 paper.qmd
Update paper.qmd

---
title: "Reproducible development environments with rix"
format:
    jss-pdf:
        keep-tex: true
    jss-html: default
author:
  - name: Bruno Rodrigues 
    affiliations:
      - name: Ministry of Research and Higher education, Luxembourg 
        department: Department of Statistics
        address: 18, Montée de la Pétrusse
        city: Luxembourg 
        country: Luxembourg 
        postal-code: 2327
      - Journal of Statistical Software
    orcid: 0000-0002-3211-3689
    email: bruno@brodrigues.co
    url: https://www.brodrigues.co 
  - name: Philipp Baumann
    affiliations:
      - Plus Affiliation
abstract: |
  In order create an analysis that is reproducible,
  it is not enough to write clean code and document it well. One
  must also make sure to clearly list all the dependencies of the analysis
  and ideally provide an easy way to install them.
  There are several tools that can be used to list dependencies and to
  make them easily installable by someone that wishes to reproduce a study,
  such as [Docker]{.proglang}, a containerization solution. This paper will
  present another tool that can be used to make projects reproducible,
  the [Nix]{.proglang} package manager, and an [R]{.proglang} package 
  called [rix]{.pkg} that makes using [Nix]{.proglang} much simpler 
  for users of [R]{.proglang}.

keywords: [reproducibility, R, Nix]
keywords-formatted: [reproducibility, "[R]{.proglang}", "[Nix]{.proglang}"]

@@

*Linked and executable code and data* refer to the compiled source code scripts,
but for an interpreted language such as [R]{.proglang}, this would rather
refer to the *computational environment*. The computational environment is all
the software required to run the analysis. Here too, we can speak of a
continuum. One could simply name and list the software used: for example, the
[R]{.proglang} programming language. Sometimes, authors have the courtesy to
also state the version of [R]{.proglang} used. Some authors go further, and also
list the packages used, and ideally with their versions as well. Authors rarely
state the operating system on which the analysis was done, even though it has
been shown that running the same analysis with the same software but on
different operating systems could lead to different results, as described in
@neupane2019. Authors also only very rarely provide instructions to install the
required tools and software in order to reproduce their studies.

@@

However, reaching this gold standard is quite costly: one needs to learn a tool
to deal with package versions, then [Docker]{.pkg} for all the other software,
including the programming language that was used. Ideally, a build automation
tool such as [Make]{.pkg} should be added to the list. It should also be
noted that these requirements are not limited to the [R]{.proglang} programming
language. A very similar approach should be taken if one uses
[Python]{.proglang}, or any other programming language, for statistical analysis
instead. Furthermore, the previous discussion also implies that for an analysis
to actually be reproducible at all, it must be made using free and open source
software, as proprietary software cannot be freely shared with others that wish
to reproduce a study, and there's no guarantee that others will have access to
the proprietary software used.

As an alternative to the above approach, we will present the [Nix]{.proglang}
package manager, which is available for all major operating systems and focuses
on installing and building software in a reproducible manner. This means that
this single tool can replace not only [Docker]{.pkg} but also [renv]{.pkg}
and even a build automation tool such as [Make]{.pkg}. To make
[Nix]{.pkg} more accessible to [R]{.proglang} programmers, we developed the
[rix]{.pkg} package which we will present in this article.

@@

For example, the [xlsx]{.pkg} [R]{.proglang} package requires the
[Java]{.proglang} programming language to be installed to successfully install.
Depending on what system one tries to install [xlsx]{.pkg}, installing
[Java]{.proglang} might not be easy, nor even possible. But with [Nix]{.pkg}, it
suffices to declare that the [xlsx]{.pkg} package is needed for the project, and
[Nix]{.proglang} figures out automatically that [Java]{.proglang} is required
and installs and configures it. It all just happens without any required
intervention from the user. In this context, [xlsx]{.pkg} can be referred to as
a component closure, and quoting @dolstra2004nix:

@@

But how does [Nix]{.pkg} achieve this? Where do these packages, or closures,
come from? When installing a package using [Nix]{.pkg}, an expression written in
the [Nix]{.proglang} programming language gets downloaded from the `nixpkgs`
Github repository and evaluated. This expression contains a so-called
*derivation*. A derivation defines a build: its dependencies, commands to build
and install the package in question, and then an output. Most of the time, a
derivation downloads source code, builds the software from the source and then
outputs a compiled binary. Derivations are extremely flexible, and by writing
one's own, it is possible to define and build an environment for a project in a
reproducible manner. The dependencies of these derivations are also defined in
other expressions, which will also get installed if needed. This is why the 
[Nix]{.pkg} manager *knows* that [Java]{.proglang} is needed for [xlsx]{.pkg}.

But why is installing software with [Nix]{.proglang} reproducible? Because the
whole set of [Nix]{.pkg} expressions is hosted on Github, it is possible to use
a specific commit of `nixpkgs` to ensure reproducibility of our project. Using a
specific commit of `nixpkgs` is called *pinning a revision (or commit)*. Pinning
a revision ensures that every package that Nix installs will always be at
exactly the same versions, regardless of when in the future the packages get
built. This is because the [Nix]{.pkgs} expressions that get downloaded when
pinning are the expressions as they were at that commit of `nixpkgs`.

While pinning revisions is essential for reproducibility, this is not the only
reason [Nix]{.pkg} is very useful for reproducibility, but the [Nix]{.pkg}
package manager itself was designed with reproducibility in mind. [Nix]{.pkg} is
a functional package manager, meaning that it uses concepts from functional
programming to build software: no-side effects or global variables are allowed,
and the output is always the same if given the same inputs, regardless of where
or when a package gets built (an important side-note is in order here: while the
functional aspect of [Nix]{.pkg} is quite useful to ensure reproducibility, it
can lead to situations where writing the required [Nix]{.proglang} expression to
build it can be much more difficult than usual: this is especially true for
packages that must download assets off the internet during installation, such as
several BioConductor packages like [musData]{.pkg} for example. But this is only
a concern for the packagers, not users of the package that simply wish to
install and use the package). [Nix]{.pkg} also allows for multiple versions (or
variants in the [Nix]{.pkg} jargon) to be installed at the same time on the same
system. Each variant of a package has a unique identifier, so that it is
possible to have two versions of [R]{.proglang} installed and use the correct
one where adequate. We refer to @dolstra2004nix for a more technical discussion
on [Nix]{.pkg}.

With Nix, it is essentially possible to replace [renv]{.pkg} and
[Docker]{.proglang} combined, or in the case of [Python]{.proglang}, replace
`requirements.txt` files which list package dependencies for [Python]{.proglang}
projects. It is also possible to build multi-language environments, containing R
and Python, a LaTeX distribution, and even install one's favorite text editor to
edit the project's source code. Essentially, one can define a complete,
project-specific, and reproducible environment and use it interactively, or
non-interactively. It will always be possible to rebuild this environment in the
future (as long as the `nixpkgs` GitHub repository is online) to run an analysis
sometime in the future.

@@

To define a variable in the [Nix]{.proglang} language, the `let` keyword is
used, so the first lines of this expression define the variables called `pkgs`
and `system_pkgs`. `pkgs` is defined as the set of packages to be used in this
derivation. In this example, the package expressions are going to be pulled from
the official `nixpkgs` Github repository, and in particular the commit starting
with `976fa336` is going to be used (it should be noted that is also possible to
use a personal fork of `nixpkgs` instead of the official repository).
`system_pkgs` lists the packages that we wish to install and use with this
environment; in this example only the [R]{.proglang} programming language will
be installed, alongside all of its dependencies, and all of the dependencies’
dependencies. Finally, the `mkShell` function is used to build a development
shell containing the set of packages defined by the variable `system_packages`
and its `shellHook` is set to `R --vanilla`: this means that when entering the
shell, [R]{.proglang} will be started with the `--vanilla` flag (thus ignoring
any startup options).

This expression can be saved in a file called `default.nix`, and the shell can
be built using the `nix-build` command on a system where [Nix]{.proglang} is
available.^[To install [Nix]{.proglang} we recommend the installer from
Determinate systems which can be found on the following link:
\url{https://determinate.systems/posts/determinate-nix-installer}.] Once
`nix-build` is done building the development environment according to the
specifications declared in the `default.nix` file, the user can enter the
development shell using `nix-shell`. The user will thus be dropped into an
interactive shell with all the packages declared in the `default.nix` file and
can now use this shell for work. This is similar to activating virtual
environments in the [Python]{.proglang} ecosystem.

Obviously, writing such an expression will be quite complicated for someone that
is unfamiliar with the [Nix]{.proglang} programming language, but the prospect
of being able to define a completely reproducible development environment in a
single text file and then use that file to build and use that environment
anywhere is quite appealing. Because we believe that lowering the cost of
writing [Nix]{.proglang} expressions would lead to a wider adoption of the
[Nix]{.pkg} package manager for reproducibility, we developed [rix]{.pkg}.
[rix]{.pkg} comes with the `rix()` function that helps generate such expressions
quite easily. [rix]{.pkg} is available on CRAN and can be installed as any other
package, but it is also possible to use it to bootstrap an [R]{.proglang}
development environment on a system where [R]{.proglang} is not available (but
[Nix]{.proglang} is). One simply needs to run this line inside a terminal:

```
nix-shell --expr "$(curl -sl https://raw.githubusercontent.com/ropensci/rix/main/inst/extdata/default.nix)"
```

to be dropped in a temporary [R]{.proglang} session where [rix]{.pkg} is
available. From there, it is possible to generate new [Nix]{.proglang}
expressions to build new environments. For example, here is how one could use
[rix]{.pkg} to generate a `default.nix` file that installs version 4.3.1 of
[R]{.proglang} and the [dplyr]{.pkg} and [chronicler]{.pkg} packages (we don't
show the generated expression here because of space):

@@

Here, users do not need to look for the right commit of `nixpkgs`: it is enough
to specify the version of [R]{.proglang} they wish to use and the right commit
will be selected (a lookup table linking versions of R and `nixpkgs` commits is
included in the package). The versions of [R]{.proglang} packages that were
current at the time will get installed in that environment. But [rix]{.pkg} is
quite flexible and makes it easy to install specific versions of packages (which
will get installed from CRAN) as well as packages hosted on Github and \LaTeX
packages:

```{r, eval = F}
rix(r_ver = "4.3.1",
    r_pkgs = c("dplyr", "chronicler", "AER@1.2-8"),
    system_pkgs = c("quarto", "git"),
    tex_pkgs = c(
          "amsmath",
          "framed",
          "fvextra",
          "environ",
          "fontawesome5",
          "orcidlink",
          "pdfcol",
          "tcolorbox",
          "tikzfill"
    ),
    git_pkgs = list(
                 list(package_name = "rix",
                   repo_url = "https://github.com/b-rodrigues/rix/",
                   branch_name = "master",
                   commit = "ea92a88ecdfc2d74bdf1dde3e441d008521b1756"),
                 list(package_name = "fusen",
                   repo_url = "https://github.com/ThinkR-open/fusen",
                   branch_name = "main",
                   commit = "d617172447d2947efb20ad6a4463742b8a5d79dc")
    ),
    ide = "rstudio",
    project_path = ".",
    overwrite = TRUE)
```

The above call to `rix()` generates an expression that installs several
[R]{.proglang} packages ([AER]{.pkg} version 1.2-8 gets installed, which is an
older version of the package. Had the user not specified the version, the
version available as of the `nixpkgs` commit would be used), but also installs
several TeXLive packages for authoring documents with \LaTeX, the development
versions of [rix]{.pkg} and [fusen]{.pkg} (downloaded from Github), and finally
the RStudio integrated development environment. It should be noted that as of
writing, RStudio is not available on macOS through `nixpkgs`, so the above
expression will only work on Windows and Linux distributions. To make the
expression work on macOS, the "rstudio" string could be replaced by "code",
which would install VS Code instead, or "other", to use any other editor such as
Emacs or Vim. You can see the resulting `default.nix` file in the @sec-appendix.

[rix]{.pkg}'s documentation also provides installation instructions for Nix, but
[rix]{.pkg} can generate Nix expression even if Nix is not installed on one's
system. This is quite useful if one wants to generate an expression and use that
expression in a CI/CD environment such as Github Actions. For example, the
repository that contains the source code to this
article^[https://github.com/b-rodrigues/rix_paper] uses Github Actions for
compilation. Each time one of the authors pushed to the master branch, a Github
Actions runner would be started, [Nix]{.proglang} would get installed on said
runner, and an environment containing the right packages was generated from a
`default.nix` file that is hosted on the same repository. The paper would then
get compiled using [Quarto]{.pkg} in this reproducible environment.

@@

@@

There are many tools to make one’s analyses reproducible, but Nix distinguishes
itself from these in that its main concern is in deploying the complete set of
packages that are *closed under the ''depends on'' relation*: Nix does not
simply install one piece of software, but its dependencies, and their
dependencies. This makes Nix extremely useful for reproducible research, but
because it solves a complex problem, it is a complex tool. With [rix]{.pkg}, we
wish to make using Nix more accessible by providing a familiar interface and
user experience to [R]{.proglang} users. Because it is possible to build
reproducible development shells using [Nix]{.proglang}, it can accomodate any
use case: running scripts, pipelines, interactive [shiny]{.pkg} applications,
and serve [plumber]{.pkg} application programming interfaces.

>>>>

---
title: "Reproducible development environments with rix"
format:
    jss-pdf:
        keep-tex: true
    jss-html: default
author:
  - name: Bruno Rodrigues 
    affiliations:
      - name: Ministry of Research and Higher education, Luxembourg 
        department: Department of Statistics
        address: 18, Montée de la Pétrusse
        city: Luxembourg 
        country: Luxembourg 
        postal-code: 2327
      - Journal of Statistical Software
    orcid: 0000-0002-3211-3689
    email: bruno@brodrigues.co
    url: https://www.brodrigues.co 
  - name: Philipp Baumann
    affiliations:
      - Plus Affiliation
abstract: |
  In order create an analysis that is reproducible,
  it is not enough to write clean code and document it well. One
  must also make sure to clearly list all the dependencies of the analysis
  and ideally provide an easy way to install them.
  There are several tools that can be used to list dependencies and to
  make them easily installable by someone that wishes to reproduce a study,
  such as [Docker]{.proglang}, a containerization solution. This paper will
  present another tool that can be used to make projects reproducible,
  the  [Nix]{.pkg} package manager, and an [R]{.proglang} package 
  called [rix]{.pkg} that makes using  [Nix]{.pkg} much simpler 
  for users of [R]{.proglang}.

keywords: [reproducibility, R, Nix]
keywords-formatted: [reproducibility, "[R]{.proglang}", " [Nix]{.pkg}"]

@@

*Linked and executable code and data* refer to the compiled source code scripts,
but for the purposes of this article, we will assume this meas the
*computational environment*. The computational environment is all
the software required to run the analysis. Here too, we can speak of a
continuum. One could simply name and list the software used: for example, the
[R]{.proglang} programming language. Sometimes, authors have the courtesy to
also state the version of [R]{.proglang} used. Some authors go further, and also
list the packages used, and ideally with their versions as well. Authors rarely
state the operating system on which the analysis was done, even though it has
been shown that running the same analysis with the same software but on
different operating systems could lead to different results, as described in
@neupane2019. Authors also only very rarely provide instructions to install the
required tools and software in order to reproduce their studies.

@@

However, reaching this gold standard is quite costly: one needs to learn a tool
to deal with package versions, then [Docker]{.pkg} for all the other software,
including the programming language that was used. Ideally, a build automation
tool such as [Make]{.pkg} should be added to the list. It should also be
noted that these requirements are not limited to the [R]{.proglang} programming
language. A very similar approach should be taken if one uses
[Python]{.proglang}, or any other programming language. 
Furthermore, the previous discussion also implies that for an analysis
to actually be reproducible at all, it must be made using free and open source
software, as proprietary software cannot be freely shared with others that wish
to reproduce a study, and there's no guarantee that others will have access to
the proprietary software used.

As an alternative to the above approach, we will present the  [Nix]{.pkg}
package manager, which is available for all major operating systems and focuses
on installing and building software in a reproducible manner. This means that
this single tool can replace not only [Docker]{.pkg} but also [renv]{.pkg}
and even a build automation tool such as [Make]{.pkg}. To make
[Nix]{.pkg} more accessible to [R]{.proglang} programmers, we developed the
[rix]{.pkg} package which we will present in this article.

@@

For example, the [xlsx]{.pkg} [R]{.proglang} package requires the
[Java]{.proglang} programming language to be installed to successfully install.
Depending on what system one tries to install [xlsx]{.pkg}, installing
[Java]{.proglang} might not be easy, nor even possible. But with  [Nix]{.pkg}, it
suffices to declare that the [xlsx]{.pkg} package is needed for the project, and
 [Nix]{.pkg} will install [Java]{.proglang} and configure it. This is possible
because the maintainers of the [R]{.proglang} language for  [Nix]{.pkg}
declared [Java]{.proglang} as a dependency of [xlsx]{.pkg}. 
From the perspective of the end-users, this all happens automatically, thanks
to the efforts of the maintainers.

[xlsx]{.pkg} can thus be referred to as a component closure, 
and quoting @dolstra2004nix:

@@

But how does  [Nix]{.pkg} achieve this? Where do these packages, or closures,
come from? When installing a package using  [Nix]{.pkg}, an expression written in
the  [Nix]{.pkg} programming language gets downloaded from the `nixpkgs`
Github repository and evaluated. `nixpkgs` is the name of the repository that 
contains the more than 100'000 pieces of software that can be installed with the 
[Nix]{.pkg} manager. Expressions contain so-called *derivations*. 
A derivation defines a build: its dependencies, commands to build
and install the package in question, and then an output. Most of the time, a
derivation downloads source code, builds the software from the source and then
outputs a compiled binary. Derivations are extremely flexible, and by writing
one's own, it is possible to define and build an environment for a project in a
reproducible manner. The dependencies of these derivations are also defined in
other expressions, which will also get installed if needed. As explained
above, the maintainers of [R]{.proglang} for  [Nix]{.pkg} declare the right
dependencies for the packages to turn them into component closures.

But why is installing software with  [Nix]{.pkg} reproducible? Because the
whole set of  [Nix]{.pkg} expressions is hosted on GitHhub, it is possible to use
a specific commit of `nixpkgs` to ensure reproducibility of our project. Using a
specific commit of `nixpkgs` is called *pinning a revision (or commit)*. Pinning
a revision ensures that every package that Nix installs will always be at
exactly the same versions, regardless of when in the future the packages get
built. This is because the [Nix]{.pkgs} expressions that get downloaded when
pinning are the expressions as they were at that commit of `nixpkgs`.

While pinning revisions is essential for reproducibility, this is not the only
reason  [Nix]{.pkg} is very useful for reproducibility, but the  [Nix]{.pkg}
package manager itself was designed with reproducibility in mind.  [Nix]{.pkg} is
a functional package manager, meaning that it uses concepts from functional
programming to build software: no-side effects or global variables are allowed,
and the output is always the same if given the same inputs, regardless of where
or when a package gets built (an important side-note is in order here: while the
functional aspect of  [Nix]{.pkg} is quite useful to ensure reproducibility, it
can lead to situations where writing the required  [Nix]{.pkg} expression to
build it can be much more difficult than usual: this is especially true for
packages that must download assets from the internet during installation, such as
several BioConductor packages like [musData]{.pkg} for example. But this is only
a concern for the packagers, not users of the package that simply wish to
install and use the package).  [Nix]{.pkg} also allows for multiple versions (or
variants in the  [Nix]{.pkg} jargon) to be installed at the same time on the same
system. Each variant of a package has a unique identifier, so that it is
possible to have two versions of [R]{.proglang} installed and use the correct
one where adequate. We refer to @dolstra2004nix for a more technical discussion
on  [Nix]{.pkg}.

With Nix, it is essentially possible to replace [renv]{.pkg} and
[Docker]{.proglang} combined, or in the case of [Python]{.proglang}, replace
`requirements.txt` files which list package dependencies for [Python]{.proglang}
projects and virtual environments. 
It is also possible to build multi-language environments, containing R
and Python, a LaTeX distribution, and even install one's favorite text editor to
edit the project's source code. Essentially, one can define a complete,
project-specific, and reproducible environment and use it interactively, or
non-interactively. It will always be possible to rebuild this environment in the
future (as long as the `nixpkgs` GitHub repository is online) to run an analysis
sometime in the future.

@@

To define a variable in the  [Nix]{.proglang} language, the `let` keyword is
used, so the first lines of this expression define the variables called `pkgs`
and `system_pkgs`. `pkgs` is defined as the set of packages to be used in this
derivation. In this example, the package expressions are going to be pulled from
the official `nixpkgs` GitHub repository, and in particular the commit starting
with `976fa336` is going to be used.
`system_pkgs` lists the packages that we wish to install and use with this
environment; in this example only the [R]{.proglang} programming language will
be installed, alongside all of its dependencies, and all of the dependencies’
dependencies. Finally, the `mkShell` function is used to build a development
shell containing the set of packages defined by the variable `system_packages`
and its `shellHook` is set to `R --vanilla`: this means that when entering the
shell, [R]{.proglang} will be started with the `--vanilla` flag (thus ignoring
any startup options).

This expression can be saved in a file called `default.nix`, and the shell can
be built using the `nix-build` command on a system where  [Nix]{.pkg} is
available.^[To install  [Nix]{.pkg} we recommend the installer from
Determinate systems which can be found on the following link:
\url{https://determinate.systems/posts/determinate-nix-installer}.] Once
`nix-build` is done building the development environment according to the
specifications declared in the `default.nix` file, the user can enter the
development shell using `nix-shell`. The user will thus be dropped into an
interactive shell with all the packages declared in the `default.nix` file and
can now use this shell for work. This is similar to activating virtual
environments in the [Python]{.proglang} ecosystem.

Obviously, writing such an expression will be quite complicated for someone that
is unfamiliar with the [Nix]{.proglang} programming language, but the prospect
of being able to define a completely reproducible development environment in a
single text file and then use that file to build and use that environment
anywhere is quite appealing. Because we believe that lowering the cost of
writing [Nix]{.proglang} expressions would lead to a wider adoption of the
[Nix]{.pkg} package manager for reproducibility, we developed [rix]{.pkg}.
[rix]{.pkg} comes with the `rix()` function that helps generate such expressions
quite easily. [rix]{.pkg} is available on CRAN and can be installed as any other
package, but it is also possible to use it to bootstrap an [R]{.proglang}
development environment on a system where [R]{.proglang} is not available (but
[Nix]{.pkg} is). One simply needs to run this line inside a terminal:

```
nix-shell -p R rPackages.rix"
```

to be dropped in a temporary [R]{.proglang} session where [rix]{.pkg} is
available. From there, it is possible to generate new  [Nix]{.pkg}
expressions to build new environments. For example, here is how one could use
[rix]{.pkg} to generate a `default.nix` file that installs version 4.3.1 of
[R]{.proglang} and the [dplyr]{.pkg} and [chronicler]{.pkg} packages (we don't
show the generated expression here because of space):

@@

[rix]{.pkg} can also generate much more complex expressions:

```{r, eval = F}
rix(r_ver = "4.3.1",
    r_pkgs = c("dplyr", "chronicler", "AER@1.2-8"),
    system_pkgs = c("quarto", "git"),
    tex_pkgs = c(
          "amsmath",
          "framed",
          "fvextra",
          "environ",
          "fontawesome5",
          "orcidlink",
          "pdfcol",
          "tcolorbox",
          "tikzfill"
    ),
    git_pkgs = list(
                 list(package_name = "rix",
                   repo_url = "https://github.com/b-rodrigues/rix/",
                   branch_name = "master",
                   commit = "ea92a88ecdfc2d74bdf1dde3e441d008521b1756"),
                 list(package_name = "fusen",
                   repo_url = "https://github.com/ThinkR-open/fusen",
                   branch_name = "main",
                   commit = "d617172447d2947efb20ad6a4463742b8a5d79dc")
    ),
    ide = "positron",
    project_path = ".",
    overwrite = TRUE)
```

The above call to `rix()` generates an expression that installs several
[R]{.proglang} packages ([AER]{.pkg} version 1.2-8 gets installed, which is an
older version of the package and also installs
several TeXLive packages for authoring documents with \LaTeX, the development
versions of [rix]{.pkg} and [fusen]{.pkg} (downloaded from Github), and finally
the Positron integrated development environment. You can see the resulting `default.nix` file in the @sec-appendix.

[rix]{.pkg}'s documentation also provides installation instructions for Nix, but
[rix]{.pkg} can generate Nix expression even if Nix is not installed on one's
system. This is quite useful if one wants to generate an expression and use that
expression in a CI/CD environment such as Github Actions. For example, the
repository that contains the source code to this
article^[https://github.com/b-rodrigues/rix_paper] uses Github Actions for
compilation. Each time one of the authors pushed to the master branch, a Github
Actions runner would be started,  [Nix]{.pkg} would get installed on said
runner, and an environment containing the right packages was generated from a
`default.nix` file that is hosted on the same repository. The paper would then
get compiled using [Quarto]{.pkg} in this reproducible environment.

@@

## rix uses a fork of Nixpkgs

Expressions generated with [rix]{.pkg} will point to a fork of `nixpkgs` instead
of the official, upstream repository.

@@

There are many tools to make one’s analyses reproducible, but Nix distinguishes
itself from these in that its main concern is in deploying the complete set of
packages that are *closed under the ''depends on'' relation*: Nix does not
simply install one piece of software, but its dependencies, and their
dependencies. This makes Nix extremely useful for reproducible research, but
because it solves a complex problem, it is a complex tool. With [rix]{.pkg}, we
wish to make using Nix more accessible by providing a familiar interface and
user experience to [R]{.proglang} users. Because it is possible to build
reproducible development shells using  [Nix]{.pkg}, it can accomodate any
use case: running scripts, pipelines, interactive [shiny]{.pkg} applications,
and serve [plumber]{.pkg} application programming interfaces.
